Generative Dynamics of Supreme Court Citations: Analysis with a New Statistical Network Model

### 2020-08-09 Tex/PA_rr/main.tex
added pa rr version

\documentclass[headsepline=true, abstracton]{scrartcl}

\usepackage[utf8]{inputenc}
%\usepackage[T1]{fontenc}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{bm}

\usepackage{natbib}

\usepackage[moderate]{savetrees}

\usepackage[table,xcdraw]{xcolor}

\usepackage{graphicx}

 \usepackage{float}

\usepackage{setspace}

\usepackage{url}

\usepackage{geometry}

@@

\renewcommand{\refname}{Bibliography}

\onehalfspacing
\setlength{\headsep}{15mm}

\thispagestyle{plain}

\title{\Large Generative Dynamics of Supreme Court Citations: \\ Analysis with a New Statistical Model}
%\author{}
 % Toggle % below to blind and unblind the manuscript
%  \author[1]{Christian Schmid \thanks{cxs5700@psu.edu}}
%   \author[2]{Ted Hsuan Yun Chen\thanks{thc126@psu.edu}}
% \author[2]{Bruce A. Desmarais\thanks{bdesmarais@psu.edu}}
% \author[1]{David R. Hunter \thanks{dhunter@stat.psu.edu}}
% \affil[2]{Department of Political Science, Pennsylvania State University}
 %\affil[1]{Department of Statistics, Pennsylvania State University}

\author{%
  Christian S. Schmid\footnote{Department of Statistics, The Pennsylvania State University, schmid@psu.edu}%
  \and Ted Hsuan Yun Chen \footnote{Department of Political Science, The Pennsylvania State University, thc126@psu.edu}%
   \and Bruce A. Desmarais \footnote{Department of Political Science, The Pennsylvania State University, bdesmarais@psu.edu}%
%  \and David R. Hunter \footnote{Department of Statistics, The Pennsylvania State University, dhunter@stata.psu.edu}%
  }

\maketitle
\begin{abstract}
\noindent The significance and influence of US Supreme Court majority opinions derive in large part from opinions' roles as precedents for future opinions. A growing body of literature seeks to understand what drives the use of opinions as precedents through the study of Supreme Court case citation patterns. We raise two limitations of existing work on Supreme Court citations. First, dyadic citations are typically aggregated to the case level before they are analyzed. Second, citations are treated as if they arise independently. We present a methodology for studying citations between Supreme Court opinions at the dyadic level, as a network, that overcomes these limitations. This methodology---the citation exponential random graph model---enables researchers to account for the effects of case characteristics and complex forms of network dependence in citation formation. We then analyze a network that includes all Supreme Court cases decided between 1950 and 2015. We find evidence for dependence processes, including reciprocity, transitivity, and popularity. The dependence effects are as substantively and statistically significant as the effects of the exogenous covariates we include in the model, indicating that models of Supreme Court citation should incorporate both the effects of case characteristics and the structure of past citations.\footnote{
This work was supported in part by NIH, 
R01 AI36664-01. NSF grants SES- 1558661, SES-1637089, SES-1619644, and CISE-1320219. We would like to thank Rachael Hinkle and Benjamin Kassow for detailed comments on an earlier draft of this paper.}
\end{abstract}

\doublespacing
 \section{Introduction}

United States Supreme Court opinions exercise authority and influence, in part, through their roles as precedents affecting future jurisprudence in the US. The findings regarding the nature of the influences of precedent on the Supreme Court have been mixed, but the balance of the literature finds that past decisions exert some form of influence on the justices' decision making \citep{knight1996norm,richards2002jurisprudential,hansford2006politics,bailey2008does,bailey2011constrained,hitt2016measuring}. Despite a considerable body of research that focuses on the way in which relevant precedents shape decision making on the Court, relatively little work has focused on understanding which past opinions are cited by an opinion. Our focus in this paper is to provide what is, to our knowledge, the first comprehensive analysis of exactly which cases are cited in an opinion. We follow an emerging body of work on legal citations, and treat the system of citations as a network \citep[e.g., ][]{harris1982structural,caldeira1988legal,fowler2007network, fowler2008authority,bommarito2009law,lupu2012precedent,pelc2014politics,ethayarajh2018rose}.

@@

\begin{figure}[htp]
\centering
\includegraphics[width = 0.95\textwidth,trim= 0cm 1cm 0cm 2cm,clip=true ]{images/citations_trans.pdf}
\caption{Illustration of transitive triangle connecting US Supreme Court opinions through citations.}
\label{fig:transitivity}
\end{figure}

@@

The second network property we consider in the context of Supreme Court citations is reciprocity. Reciprocity (also referred to as mutuality) is the tendency for node B to send a tie to node A in response to or coordination with A sending a tie to B \citep{garlaschelli2004patterns}. It is typically not possible for reciprocated ties to form in legal opinions. Most citations reference past opinions that were issued before the citing opinion's case was even argued before the Court.  However, opinions written within the same Supreme Court term are often drafted in tandem, and can cite each other reciprocally.  Opinion A citing opinion B within the same term represents a signal that opinion B is relevant to the legal reasoning underpinning opinion A. Unlike the citations themselves, the applicability of legal rules or lines of reasoning across cases is not directed---if A is relevant to B, B is highly likely to be relevant to A. We expect opinions written within the same term to exhibit a high degree of reciprocity. Below we provide two example passages from opinions that illustrate the phenomenon of within-term reciprocity.

\begin{figure}[ht]
\centering
\includegraphics[width = 0.95\textwidth,trim= 0cm 1cm 0cm 2cm,clip=true ]{images/citations_recip.pdf}
\caption{Illustration of a reciprocal tie between two US Supreme Court opinions.}
\label{fig:reciprocity}
\end{figure}

The first case in our example reciprocal dyad is a passage from Western Air Lines v. Criswell (472 U.S. 400, 1985)---a case considering mandatory retirement in the context of age discrimination laws. The second case in the dyad, Johnson v. Mayor of Baltimore (472 U.S. 353, 1985) is another case considering whether mandatory retirement violates the Age Discrimination in Employment Act. The mutual edge connecting these two cases is visualized in Figure \ref{fig:reciprocity}. These cases addressed very similar legal questions, which increased the likelihood that they would inform each other, and the opinions were written within the same term, which made it possible for them to cite each other. 
\begin{quotation}
{\em From Western Air Lines:} On a more specific level, Western argues that flight engineers must meet the same stringent qualifications as pilots, and that it was therefore quite logical to extend to flight engineers the FAA's age 60 retirement rule for pilots. Although the FAA's rule for pilots, adopted for safety reasons, is relevant evidence in the airline's BFOQ defense, it is not to be accorded conclusive weight. Johnson v. Mayor and City Council of Baltimore, ante at 472 U. S. 370-371. The extent to which the rule is probative varies with the weight of the evidence supporting its safety rationale and "the congruity between the . . . occupations at issue." Ante at 472 U. S. 371. In this case, the evidence clearly established that the FAA, Western, and other airlines all recognized that the qualifications for a flight engineer were less rigorous than those required for a pilot.
\\~\\
{\em From Johnson:} The city, supported by several amici, argues for affirmance nonetheless. It asserts first that the federal civil service statute is not just a federal retirement provision unrelated to the ADEA, but in fact establishes age as a BFOQ for federal firefighters based on factors that properly go into that determination under the ADEA, see Western Air Lines, Inc. v. Criswell, post p. 472 U. S. 400. Second, the city asserts, a congressional finding that age is a BFOQ for a certain occupation is dispositive of that determination with respect to nonfederal employees in that occupation. 
\end{quotation} % https://supreme.justia.com/cases/federal/us/472/400/case.html \url{https://supreme.justia.com/cases/federal/us/472/353/case.html#370}

@@

\begin{figure}[ht]
\centering
\includegraphics[width = 0.95\textwidth,trim= 0cm 3cm 6cm 8cm,clip=true ]{images/citations_pop.pdf}
\caption{Illustration of ties sent to a landmark Supreme Court opinion via citations.}
\label{fig:popularity}
\end{figure}

These three properties---transitivity, reciprocity, and popularity form the core of our network theory of U.S. Supreme Court citations. We also seek to model the effects of covariates (i.e., case features) on citation formation.  In the next section, we introduce a modeling framework that can incorporate all of these effects into a single model, and adapt to the peculiar structure of citation matrices.

\section{The Citation Exponential Random Graph Model}
We develop a methodology that can be used to jointly test for the effects of covariates on citations---as have been studied in prior research, and test for dependence effects, as we hypothesize above. To accomplish this, we extend a model that has been developed to jointly represent covariate and dependence effects in network data---the exponential random graph model (ERGM) \citep[e.g.,][]{duque2018recognizing,osei2018elite}.  %The ERGM, as it is currently implemented \citep{leifeld2017temporal,  tergm}, is designed to model longitudinal network data in which ties among a relatively stable set of nodes are modeled at multiple time points (e.g., the countries that are and are not defensively allied at time $t$, the legislators who have and have not cosponsored each others' legislation in the current session).  
The network structures for which ERGMs are currently designed are insufficient to account for the structure of citation networks.

@@

\begin{figure}[H]
\centering
\includegraphics[width = 0.95\textwidth,trim= 0cm 0cm 0cm 0cm,clip=true  ]{images/daggish.pdf}
\caption{Illustration of temporal structure of the Supreme Court Citation Network. $C_{\leq t}$ is the entire set of citations (and non-citations) on which citations and non-citations at time $t$ (i.e., $C_t$) depend. $C_t$ are conditioned on the citations and non-citations established before time $t$ (i.e., $C_{< t}$). The shaded small squares are hypothetical observed citations, and the white small squares are citations that could have been observed but were not. The regions of the citation matrix that are represented by large white rectangles are citations that could not have been observed since the citing case would have been decided in a term that preceded the term of the cited case. The citing case ID is given in the row and the prospective cited case is given in the column. }
\label{fig:ctergm}
\end{figure}

@@

%We should note that there is nothing about the c-ERGM (or any ERGM family model, for that matter) that permits the identification of causal effects of one tie on the formation of another. For example, if we find a positive popularity effect, the term for that dynamic would indicate that there is an increase in the log odds that case $i$ cites case $j$ with each other case that has cited case $j$.  That finding could be driven by one of two (or some combination of both) dynamics. First, it could be the case that a large volume of citations to a case in past opinions draws the attention of the justices to the ``popular'' case---as in the Baker v. Carr passage cited above. Second, it could be that we have omitted a case-level variable that is strongly associated with the number of citations accumulated by a case, and the number of citations in past cases is operating as a proxy for this unmeasured case characteristic. We cannot statistically differentiate between these two dynamics. However, in both of these conditions, and analogous conditions related to other dependence properties, citations (or non-citations) to a case are not independent conditional on the other variables, and the c-ERGM can be specified to account for the dependence in the data. Estimation of the c-ERGM parameters, which is done with a computationally-intensive, simulation-based approach, is presented in detail in the appendix.

%Methods of Monte Carlo approximation are well established in the ERGM literature, \citep{hunter2006inference,van2009framework,hummel2012improving}, but with a number of nodes on the order of $10,000$, even the methods of Monte Carlo MLE are computationally expensive \citep{schmid2017exponential}. However, because the network is large, we can rely on a fast estimation method---maximum pseudo-likelihood estimation (MPLE), which is consistent for discrete exponential family models \citep{nguyen2017near}, and has long been used with ERGM \citep{strauss1990pseudolikelihood}. The MPLE is defined as maximizing the log product of the conditional probability of each citation (and non-citation), conditional on the other elements of the observed citation network. The joint probability of all citations is replaced by the product over conditional probabilities, which, as we demonstrated above, assume a logit form.

%One challenge that arises with the use of MPLE is that the usual estimate of the asymptotic covariance matrix (i.e., inverting the negated Hessian of the objective function at the maximum), results in standard errors that are biased downward. This leads to a high Type-1 error rate in hypothesis testing with MPLE if conventional standard errors are used \citep{van2009framework}. A form of non-parametric cluster bootstrap results in confidence intervals with correct coverage probability with MPLE \citep{desmarais2012statistical,desmarais2010consistent}. In the case of c-TERGM, the clusters are defined by $t$---the time points in which the cases are decided, and $T$ time-specific citation matrices $C_{t}$ are re-sampled with replacement to form a non-parametric cluster bootstrap sample. Confidence intervals are defined by the quantiles in the empirical distributions of estimates that result from re-estimating the MPLE on non-parametric cluster bootstrap samples.

@@

\begin{table}[htp]
\centering
\begin{tabular}{|
>{\columncolor[HTML]{C0C0C0}}l |l|l|l|}
\hline
{\color[HTML]{333333} } & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Terms} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Total Number Cases} & \cellcolor[HTML]{C0C0C0}{\color[HTML]{333333} Cases/Term} \\ \hline
CE Hughes*              & 1937 - 1941                                          & 628                                                               & 125.6                                                     \\ \hline
HF Stone                & 1942 - 1946                                          & 756                                                               & 151.2                                                     \\ \hline
FM Vinson               & 1946 - 1953                                          & 789                                                               & 98.63                                                      \\ \hline
E Warren                & 1954 - 1969                                          & 2149                                                              & 126.41                                                     \\ \hline
WE Burger               & 1970 - 1986                                          & 2805                                                             & 155.83                                                     \\ \hline
W Rehnquist            & 1987 - 2005                                          & 2022                                                              & 106.42                                                      \\ \hline
J Roberts **           & 2006 - 2015                                          & 871                                                              & 87.1                                           \\ \hline
\end{tabular}
\caption[caption]{For the time range of interest (1937 - 2015) this table displays the chief justices, the time range they served as chief justice, the number of cases in their time range as well as the average number of cases per year.\\\hspace{\textwidth} * CE Hughes served as chief justice from 1930 - 1941. \\\hspace{\textwidth} ** J Roberts still serves as chief justice (retrieved 5/2020).}
\label{tab:chiefs}
\end{table}

The degree distributions indicate that there is a long tail to both the number of citations sent and received. These long-tailed (i.e., high kurtosis) distributions provide preliminary evidence of substantial heterogeneity in the features that drive citations to and from cases \citep{strogatz2001exploring}. Figure \ref{fig:networkviz} displays the citation network network from 1937--2001. We see here that the densest rates of tie formation tend to be between consecutive courts (e.g., the Stone Court is much more tightly tied to the Hughes Court than the Rehnquist Court is to the Hughes Court). This pattern lends preliminary/descriptive support to the hypothesis that the rate of citations to a case decrease over time.

\begin{figure}[H]
\includegraphics[scale=0.5]{degree_distribution}
\caption{The in- and outdegree distribution of the Supreme Court Citation Network from 1937 - 2001. There are cases with an indegree  $>$50, but they are not captured in this figure.}
 \label{degree_dist}
\vspace{-.25cm}
\end{figure}

\begin{figure}[H]
\includegraphics[scale=0.35,clip=true,trim=.5cm 0cm 0cm 2cm]{images/citations1}
\caption{ Supreme Court Citation Network, 1937-2001. Nodes are Supreme Court cases, with size based on incoming citations. Salient cases ({\em Oxford Guide}) are in yellow. Each network layer contains cases decided under a different chief justice. Because our data is temporally bound, cases for the Hughes and Rehnquist courts are not complete. }
 \label{fig:networkviz}
\vspace{-.25cm}
\end{figure}

@@

We include one covariate that accounts for the rate at which cases that fall under the same issue area cite each other. The variable $X^{ij}$ indicates whether cases $i$ and $j$ share the same issue area, e.g., $X^{ij}=1$ if cases $i$ and $j$ are classified into the same issue area, and $0$ otherwise. The Issue area data comes from the Supreme Court Database (SCDB) \citep{spaeth2014supreme}. We include these variables because \citet{cross2010determinants} finds that the number of citations to Supreme Court opinions depends heavily on the issue area of the case. Table \ref{issue_area_coding} reports the issue areas covered in our data.

@@

\begin{table}[]
\centering
\begin{tabular}{llll}
1 & Criminal Procedure & 8  & Economic Activity    \\
2 & Civil Rights       & 9  & Judicial Power       \\
3 & First Amendment    & 10 & Federalism           \\
4 & Due Process        & 11 & Interstate Relations \\
5 & Privacy            & 12 & Federal Taxation     \\
6 & Attorneys          & 13 & Miscellaneous        \\
7 & Unions             & 14 & Private Action      
\end{tabular}
\caption{Assigned numbers for the variable \textit{Issue Area}. This information originates from the Supreme Court Database code book.}
\label{issue_area_coding}
\end{table}

We model the way in which citations to a case depend upon the age of a case. For this we use a second-order polynomial in which one covariate $X^{ij}$ is defined as the age of case $j$ at the time that case $i$ is decided, and another term in which $X^{ij}$ is the squared age of case $j$ at the time that case $i$ is decided. We include these covariates because \citet{black2013citation} find that the number of citations to a Supreme Court case over time depends significantly on the age of the case, characterized by a sharp drop off and leveling out with age.

\citet{benjamin2012standing} study the propensity for cases to be overruled and cited in other negative ways. They find that cases with majority coalitions that are large and ideologically broad are less likely to be cited negatively. In our data we do not differentiate between negative and positive citations, but since the overwhelming majority of citations are positive, we hypothesize that the effects they found will be reversed in our analysis. We include one covariate in which $X^{ij}$ is the number of justices in the majority coalition for case $j$. We also include another covariate in which $X^{ij}$ is the absolute difference between the maximum and minimum ideal points of the justices in the majority coalition for case $j$. We expect both of these covariates to have positive effects.

@@

%The log odds interpretation of the triangle statistic is that  $\theta$ gives the change in the log odds of case $i$ citing case $j$ associated  with each additional triangle that would be created by case $i$ citing case $j$ (e.g., if opinions in cases $i$ and $j$ cite one common case $k$, the log odds that case $i$ cites case $j$ change by theta relative to the configuration in which cases $i$ and $j$ do not cite any of the same cases).

 % \subsubsection{Dynamics in effects}

% For the current model this section is not relevant anymore

@@

  \begin{figure}[H]
\includegraphics[width=15cm]{SCC_AIC_BIC}
\caption{AIC and BIC for the full and the independent model for the time frame 1950-2015 }
 \label{AIC_BIC}
\vspace{-.25cm}
\end{figure}

@@

%In our prediction experiment we randomly split the directed dyads into an 80\% training set and a 20\% test set. The parameters of the model are estimated using the directed dyads in the training set, and the parameters are used to form the conditional probability of a tie for all of the directed dyads in the test set. Directed dyads for which the conditional probability of a tie exceeds 0.5 are predicted to be citations. The experiment is run with the full model, and with a model that excludes all of the dependence terms (i.e., all terms involving reciprocity, in and out stars, and triangles)---the independent dyads model. We run this experiment for 10 iterations. Predictive performance is evaluated with three common and related measures---precision (i.e., the proportion of predicted citations that are actually citations), recall (the proportion of actual citations that are predicted to be citations), and the F1 score (the harmonic mean of precision and recall) \citep[see, e.g.,][for discussion of these measures]{makhoul1999performance}. All three measures are bounded between 0 and 1, with higher scores indicating better performance.

%\begin{table}[H]
%\centering

%\begin{tabular}{rllll}
%\hline \hline
%& \multicolumn{2}{l}{Independent Model } & \multicolumn{2}{l}{Full Model} \\ \hline 
% & mean & range & mean & range \\ 
%  \hline
%precision & 0.6007 & (0.5934, 0.6074) & 0.8666 & (0.8647, 0.8701) \\ 
%  recall & 0.1522 & (0.1495, 0.1549) & 0.5969 & (0.5945, 0.6019) \\ 
%  F1 score & 0.2429 & (0.2389, 0.2462) & 0.7069 & (0.7048, 0.7107) \\ 
%   \hline \hline
%\end{tabular}
%\caption{The predictive performance of the directed dyadic models with linear trends, over ten 80/20 train/test splits.}
%\label{tab:prediction_linear}

%\end{table}

%We see from Table \ref{tab:prediction_linear} that, based on all three measures, the predictive performance of the model improves dramatically from adding the network dependence terms. The recall of the full model is particularly impressive, indicating that it recovers over half of the actual citations in the test set. This provides clear evidence that the full model, which includes covariates and network dependence terms, represents a more accurate and complete model of the process of citation formation in US Supreme Court opinions.

>>>>

\documentclass{cup-pan}
\usepackage[utf8]{inputenc}
\usepackage{blindtext}
\usepackage{bm}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{geometry}

\title{Generative Dynamics of Supreme Court Citations: Analysis with a New Statistical Network Model}

\author[1]{Christian S. Schmid}
\author[2]{Ted Hsuan Yun Chen}
\author[3]{Bruce A. Desmarais}

\affil[1]{Department of Statistics, The Pennsylvania State University, University Park, PA. Email: \url{schmid@psu.edu}}
\affil[2]{Faculty of Social Sciences, University of Helsinki, Finland. Email: \url{ted.hsuanyun.chen@gmail.com}}
\affil[3]{Department of Political Science, The Pennsylvania State University, University Park, PA. Email: \url{bdesmarais@psu.edu}}

%% Corresponding author
\corrauthor{Christian S. Schmid}

%% Abbreviated author list for the running footer
\runningauthor{Schmid et al.}

\addbibresource{bib.bib}

@@

\maketitle

\begin{abstract}
\noindent The significance and influence of US Supreme Court majority opinions derive in large part from opinions' roles as precedents for future opinions. A growing body of literature seeks to understand what drives the use of opinions as precedents through the study of Supreme Court case citation patterns. We raise two limitations of existing work on Supreme Court citations. First, dyadic citations are typically aggregated to the case level before they are analyzed. Second, citations are treated as if they arise independently. We present a methodology for studying citations between Supreme Court opinions at the dyadic level, as a network, that overcomes these limitations. This methodology---the citation exponential random graph model---enables researchers to account for the effects of case characteristics and complex forms of network dependence in citation formation. We then analyze a network that includes all Supreme Court cases decided between 1950 and 2015. We find evidence for dependence processes, including reciprocity, transitivity, and popularity. The dependence effects are as substantively and statistically significant as the effects of the exogenous covariates we include in the model, indicating that models of Supreme Court citation should incorporate both the effects of case characteristics and the structure of past citations.\footnote{
This work was supported in part by NSF grants SES-1558661, SES-1637089, SES-1619644, and CISE-1320219. }

\keywords{US Supreme Court, majority opinions, citations, network, ERGM.}
\end{abstract}

\section{Introduction}
\label{sec:overview}

United States Supreme Court opinions exercise authority and influence, in part, through their roles as precedents affecting future jurisprudence in the US. The findings regarding the nature of the influences of precedent on the Supreme Court have been mixed, but the balance of the literature finds that past decisions exert some form of influence on the justices' decision making \citep{knight1996norm,gillman2001s,richards2002jurisprudential,hansford2006politics,bailey2008does,bailey2011constrained,pang2012endogenous}. Despite a considerable body of research that focuses on how precedents shape decision making on the Court, relatively little work has focused on understanding which past opinions are cited by an opinion. Our focus in this paper is to provide what is, to our knowledge, the first comprehensive analysis of exactly which cases are cited in an opinion. We follow an emerging body of work on legal citations, and treat the system of citations as a network \citep[e.g., ][]{harris1982structural,caldeira1988legal,fowler2007network, fowler2008authority,bommarito2009law,lupu2012precedent,pelc2014politics,ethayarajh2018rose}.

@@

\begin{figure}[bt]
\centering
\includegraphics[width = 0.49\textwidth,trim= 2cm 2cm 3cm 2cm,clip=true ]{citations_trans.pdf}\hspace{0.2cm}
\includegraphics[width = 0.49\textwidth,trim= 2cm 2cm 3cm 2cm,clip=true ]{citations_recip.pdf}
\caption{Illustrations of transitive triangle connecting US Supreme Court opinions through citations (left) and a reciprocal tie between two US Supreme Court opinions (right).}
\label{fig:transitivity}
\end{figure}

@@

% \begin{figure}[bt]
% \centering
% \includegraphics[width = 0.85\textwidth,trim= 0cm 1cm 0cm 2cm,clip=true ]{citations_recip.pdf}
% \caption{Illustration of a reciprocal tie between two US Supreme Court opinions.}
% \label{fig:reciprocity}
% \end{figure}

The first case in our example reciprocal dyad is a passage from Western Air Lines v. Criswell (472 U.S. 400, 1985)---a case considering mandatory retirement in the context of age discrimination laws. The second case in the dyad, Johnson v. Mayor of Baltimore (472 U.S. 353, 1985) is another case considering whether mandatory retirement violates the Age Discrimination in Employment Act. The mutual edge connecting these two cases is visualized in Figure \ref{fig:transitivity}. These cases addressed very similar legal questions, which increased the likelihood that they would inform each other, and the opinions were written within the same term, which made it possible for them to cite each other. 
\begin{quotation}
{\em From Western Air Lines:} On a more specific level, Western argues that flight engineers must meet the same stringent qualifications as pilots, and that it was therefore quite logical to extend to flight engineers the FAA's age 60 retirement rule for pilots. Although the FAA's rule for pilots, adopted for safety reasons, is relevant evidence in the airline's BFOQ defense, it is not to be accorded conclusive weight. Johnson v. Mayor and City Council of Baltimore, ante at 472 U. S. 370-371. The extent to which the rule is probative varies with the weight of the evidence supporting its safety rationale and "the congruity between the . . . occupations at issue." Ante at 472 U. S. 371. In this case, the evidence clearly established that the FAA, Western, and other airlines all recognized that the qualifications for a flight engineer were less rigorous than those required for a pilot.
\\~\\
{\em From Johnson:} The city, supported by several amici, argues for affirmance nonetheless. It asserts first that the federal civil service statute is not just a federal retirement provision unrelated to the ADEA, but in fact establishes age as a BFOQ for federal firefighters based on factors that properly go into that determination under the ADEA, see Western Air Lines, Inc. v. Criswell, post p. 472 U. S. 400. Second, the city asserts, a congressional finding that age is a BFOQ for a certain occupation is dispositive of that determination with respect to nonfederal employees in that occupation. 
\end{quotation} % https://supreme.justia.com/cases/federal/us/472/400/case.html \url{https://supreme.justia.com/cases/federal/us/472/353/case.html#370}

@@

\begin{figure}[bt]
\centering
\includegraphics[width = 0.8\textwidth,trim= 1cm 4cm 1cm 3cm,clip=true ]{citations_pop.pdf}
\caption{Illustration of ties sent to a landmark Supreme Court opinion via citations.}
\label{fig:popularity}
\end{figure}

These three properties---transitivity, reciprocity, and popularity form the core of our network theory of U.S. Supreme Court citations. We also seek to model the effects of covariates (i.e., case features) on citation formation.  Next we introduce a modeling framework that can incorporate all of these effects into a single model, and adapt to the structural constraints of citation networks.

\section{The Citation Exponential Random Graph Model}
We develop a methodology that can be used to jointly test for the effects of covariates on citations---as have been studied in prior research, and test for dependence effects, as we hypothesize above. To accomplish this, we extend a model that has been developed to jointly represent covariate and dependence effects in network data, and has seen extensive application in recent political networks research---the exponential random graph model (ERGM) \citep[e.g.,][]{box2014evolution,duque2018recognizing,osei2018elite}.  %The ERGM, as it is currently implemented \citep{leifeld2017temporal,  tergm}, is designed to model longitudinal network data in which ties among a relatively stable set of nodes are modeled at multiple time points (e.g., the countries that are and are not defensively allied at time $t$, the legislators who have and have not cosponsored each others' legislation in the current session).  
The network structures for which ERGMs are currently designed are insufficient to account for the structure of citation networks.

@@

\begin{figure}[bt]
\centering
\includegraphics[width = 0.85\textwidth ]{daggish.pdf}
\caption{Illustration of temporal structure of the Supreme Court Citation Network. $C_{\leq t}$ is the entire set of citations (and non-citations) on which citations and non-citations at time $t$ (i.e., $C_t$) depend. $C_t$ are conditioned on the citations and non-citations established before time $t$ (i.e., $C_{< t}$). The shaded small squares are hypothetical observed citations, and the white small squares are citations that could have been observed but were not. The regions of the citation matrix that are represented by large white rectangles are citations that could not have been observed since the citing case would have been decided in a term that preceded the term of the cited case. The citing case ID is given in the row and the prospective cited case is given in the column. }
\label{fig:ctergm}
\end{figure}

@@

@@

\begin{table}[bt]
\caption{For the time range of interest (1937 - 2015) this table displays the chief justices, the time range they served as chief justice, the number of cases in their time range as well as the average number of cases per year.\\\hspace{\textwidth} * CE Hughes served as chief justice from 1930 - 1941. \\\hspace{\textwidth} ** J Roberts still serves as chief justice (retrieved 5/2020).}
\label{tab:chiefs}
\centering
\begin{tabular}{l l l l r}
\headrow \thead{Chief Justice} & \thead{Terms} & \thead{Total Number Cases} & \thead{Cases/Term}  \\
CE Hughes*              & 1937 - 1941                                          & 628                                                               & 125.6                                                     \\ 
HF Stone                & 1942 - 1946                                          & 756                                                               & 151.2                                                     \\ 
FM Vinson               & 1946 - 1953                                          & 789                                                               & 98.63                                                      \\ 
E Warren                & 1954 - 1969                                          & 2149                                                              & 126.41                                                     \\ 
WE Burger               & 1970 - 1986                                          & 2805                                                             & 155.83                                                     \\ 
W Rehnquist            & 1987 - 2005                                          & 2022                                                              & 106.42                                                      \\ 
J Roberts **           & 2006 - 2015                                          & 871                                                              & 87.1                                           \\ 
\end{tabular}

\end{table}

The degree distributions indicate that there is a long tail to both the number of citations sent and received. These long-tailed (i.e., high kurtosis) distributions provide preliminary evidence of substantial heterogeneity in the features that drive citations to and from cases \citep{strogatz2001exploring}. Figure \ref{degree_dist} displays the citation network network from 1937--2015. We see here that the densest rates of tie formation tend to be between consecutive courts (e.g., the Stone Court is much more tightly tied to the Hughes Court than the Rehnquist Court is to the Hughes Court). This pattern lends preliminary support to the hypothesis that the rate of citations to a case decrease over time.

\begin{figure}[bt]
\centering
\begin{minipage}{.37\linewidth}
\centering
\includegraphics[width = 0.99\textwidth]{degree_distribution.pdf}
\includegraphics[width = 0.8\textwidth]{data_mat.png}%
\end{minipage}%
\begin{minipage}{.63\linewidth}
\includegraphics[width = \textwidth]{courts_layer.png}
% \includegraphics[width = \textwidth,clip=true,trim=0.5cm 3.75cm 0.75cm 0.5cm ]{courts_layer_updated.pdf}
\end{minipage}
\caption{Supreme Court Citation Network, 1937-2015. Network visualization on the right. Nodes are Supreme Court cases, color-coded based on the chief justice presiding over the court. On the top left is the in- and outdegree distribution of the network. There are cases with an in- or outdegree  $>$50, but they are not captured in this figure. The bottom left shows the citation data in adjacency matrix format following \autoref{fig:ctergm}.}
 \label{degree_dist}
\end{figure}

% \begin{figure}[bt]
% \centering
% \includegraphics[scale = 0.45 ]{degree_distribution.pdf}
% \caption{The in- and outdegree distribution of the Supreme Court Citation Network from 1937 - 2015. There are cases with an in- or outdegree  $>$50, but they are not captured in this figure.}
%  \label{degree_dist}
% \end{figure}

% \begin{figure}[h]
% \centering
% \includegraphics[scale=0.20,clip=true,trim=.5cm 2cm 0cm 0cm ]{courts_layer_updated.pdf}
% \caption{ Supreme Court Citation Network, 1937-2015. Nodes are Supreme Court cases, color-coded based on the chief justice presiding over the court. Each network layer contains cases decided under a different chief justice. Because our data is temporally bound, cases for the Hughes and Roberts courts are not complete. }
%  \label{fig:networkviz}\end{figure}

@@

We include one covariate that accounts for the rate at which cases that fall under the same issue area cite each other. The variable $X_{ij}$ indicates whether cases $i$ and $j$ share the same issue area, e.g., $X_{ij}=1$ if cases $i$ and $j$ are classified into the same issue area, and $0$ otherwise. The Issue area data comes from the Supreme Court Database (SCDB) \citep{spaeth2014supreme}. We include these variables because \citet{cross2010determinants} finds that the number of citations to Supreme Court opinions depends heavily on the issue area of the case. Table \ref{issue_area_coding} reports the issue areas covered in our data.

@@

\begin{table}[bt]
\caption{Assigned numbers for the variable \textit{Issue Area}. This information originates from the Supreme Court Database code book.}
\label{issue_area_coding}
\centering
\begin{tabular}{llll}
1 & Criminal Procedure & 8  & Economic Activity    \\
2 & Civil Rights       & 9  & Judicial Power       \\
3 & First Amendment    & 10 & Federalism           \\
4 & Due Process        & 11 & Interstate Relations \\
5 & Privacy            & 12 & Federal Taxation     \\
6 & Attorneys          & 13 & Miscellaneous        \\
7 & Unions             & 14 & Private Action      
\end{tabular}
\end{table}

We model the way in which citations to a case depend upon the age of a case. For this we use a second-order polynomial in which one covariate $X_{ij}$ is defined as the age of case $j$ at the time that case $i$ is decided, and another term in which $X_{ij}$ is the squared age of case $j$ at the time that case $i$ is decided. We include these covariates because \citet{black2013citation} find that the number of citations to a Supreme Court case over time depends significantly on the age of the case, characterized by a sharp drop off and leveling out with age.

\citet{benjamin2012standing} study the propensity for cases to be overruled and cited in other negative ways. They find that cases with majority coalitions that are large and ideologically broad are less likely to be cited negatively. In our data we do not differentiate between negative and positive citations, but since the overwhelming majority of citations are positive, we hypothesize that the effects they found will be reversed in our analysis. We include one covariate in which $X_{ij}$ is the number of justices in the majority coalition for case $j$. We also include another covariate in which $X_{ij}$ is the absolute difference between the maximum and minimum ideal points of the justices in the majority coalition for case $j$. We expect these covariates to have positive effects.

@@

\begin{figure}[bt]
\centering
\includegraphics[width=10cm ]{SCC_AIC_BIC.pdf}
\caption{AIC and BIC for the full and the independent model for the time frame 1950-2015 }
 \label{AIC_BIC}
\end{figure}

@@

@@
